### 2016-02-10 reviews/answer_to_revieweres.txt

ANSWER: this was done, and modified in the text

> Line 284: Figure S1 does not contain information at the ‘per individual’ level

ANSWER: updated the text.

> Line 304: How sensitive are the results to the imposed quality filters?

ANSWER: not very sensitive. Quality filters have a direct effect on the elaboration steps following imputation (GWAS in particular) but little impact on imputation itself. Only very relaxed filtering would change imputation performances, since a lot of noisy data point would be included. This scenario is however very far from standard practice and of little interest for the current work.

@@

> Line 313/Figure 3: Suggest to maintain consistency in y-axis scale, i.e., the plots for Minor homozygous accuracy should range from 0 to 100% 
as well.

ANSWER: Reviewer 1 had the opposite suggestion, which we followed. See answer above.

> Line 324: the flat response seems to hold for the studied range (0.01 – 0.20). Were higher values of missing rates studied?  It will be interesting to study whether there will be a critical point and whether these are different for the different genotypic classes.

ANSWER: the cited range (0.01 - 0.20) is of injected missing data points, and it overlaps the other range (0.10 - 0.70) of naturally occurring missing data points. This brings the maximum accepted level (worst case scenario) to 0.2 + 0.7 = 0.9, which seems a good place where to stop. Please keep in mind that published works using GBS data usually filter on the 0.3-0.5 range, so we think that our choice abundantly covers the current practice.

> Line 347: Rather extensive reporting on computation time which seems not justified as the used methods were used as these were available to the authors, i.e.., there was not dedicated efforts to speed up the implemented methods.

> Line 379-381: Where was this negative impact shown?  In this study or other studies – please include references.

ANSWER: a reference to Rutkoski et al was added. We tried to explain better the difference between SNP array (~5% missing) and GBS data (~50% missing)

> Line 401-407: This is still rather descriptive without any postulation whether any of the methods are less or more sensitive to MAF and/or balance.  The superior results of Beagle in rice are mostly driven by utilizing map/assembly information which is ignored by other methods.

ANSWER We updated the section, improving the description and commenting the results of an extra method that uses marker information.

> Line 440: The reasoning on factor ii) is not clear – would the results for rice be less accurate when analysing the genome at once? Also, why would it simplify the imputation process?

ANSWER: First, the per-chromosome analysis simplifies calculations, and allow for straightforward parallelization. Additionally, the results with rice as a whole are less accurate when markers are shuffled [add numbers/supplementary Table/Figure?]; this is because the LD structure is better preserved when reshuffling markers.

> Line 443: Did you explore the population structure of the three datasets? Would presence/absence of (severe) population structure affect imputation accuracy?

Answer: i) no structure in alfalfa, presence of structure in rice (indica/tropical/temperate); ii) implicitly exploited by most imputation methods (e.g. KNN --> distance based on genetic relatedness); iii) not likely to affect much imputation accuracy; iv) this is not GWAS: the interest is not in estimating the effect of the SNP (which may be inflated if other factors are not included in the model), but in "prediction/classification" of the missing genotype to be imputed;

> Line 450: It is too easy to state that there is additional complexity to make use of a reference genome of a related species. It would be of HIGH relevance to explore this as it may be applied in many species that lack a reference genome themselves.

ANSWER: we added the study using Medicago Truncatula.

### 2016-02-10 inputs/discussion.tex
phrasing

In rice, where a reference genome is available, we could assess the effect of using ordered vs. unordered markers for imputation. For alfalfa there is no reference genome sequenced yet. The genome of the close relative diploid \emph{Medicago Truncatula} is available \cite{young_medicago_2011} and can in principle used. However, while the two genomes show high synteny \cite{li_saturated_2014}, aligning on a different species (and with different ploidity) comes at the price of discarding those markers that do not align. \textcolor{red}{HOW MANY?}\\

Among the imputation methods used in this study, Beagle and FILLIN are the only ones that make use of marker information, and were therefore tested with ordered and randomly shuffled markers. In rice dataset with ordered markers Beagle showed astounding resilience to high missingness, with imputation accuracy in the minority class $>$95\% even in the most extreme scenarios (70\% allowed, 20\% artificial missing genotypes).
Beagle worked substantially worse when the marker position was randomly reshuffled, with imputation accuracy in the minority class ranging from 75\% (10\% allowed, 1\% artificial missing rates) to almost 50\% (70\% allowed, 20\% artificial missing rates). When tested with alfalfa Beagle largerly overestimated by the majority class and had very poor performances on heterozygous class. Shuffling marker positions further worsened the accuracy, and almost all missing data point assigned to the majority class.\\
FILLIN performances on rice follow the same pattern and worsen substantially when markers are shuffled. This result confirmed that imputation methods specifically designed for ordered markers are not an option for species lacking a reference genome.

>>>>

In rice, where a reference genome is available, we could assess the effect of using ordered vs. unordered markers for imputation. For alfalfa there is no reference genome sequenced yet. The genome of the close relative diploid \emph{Medicago Truncatula} is available \cite{young_medicago_2011} and can in principle used. However, while the two genomes show high synteny \cite{li_saturated_2014}, aligning on a different species (and with different ploidity) comes at the price of discarding those markers that do not align. In our case only 57.54\% (Alfalfa-Med) and 57.86\% (Alfalfa-PV) aligned on the \emph{M. Truncatula} genome, compared to the 88.66\% of rice markers aligning on \emph{O. Sativa} genome, a loss of markers that could or could not be acceptable, depending on the specific application requirements. \\
Among the imputation methods used in this study, Beagle and FILLIN are the only ones that make use of marker information, and were therefore tested with ordered and randomly shuffled markers. In rice dataset with ordered markers Beagle showed astounding resilience to high missingness, with imputation accuracy in the minority class $>$95\% even in the most extreme scenarios (70\% allowed, 20\% artificial missing genotypes).
Beagle worked substantially worse when the marker position was randomly reshuffled, with imputation accuracy in the minority class ranging from 75\% (10\% allowed, 1\% artificial missing rates) to almost 50\% (70\% allowed, 20\% artificial missing rates). When tested with alfalfa Beagle largerly overestimated by the majority class and had very poor performances on heterozygous class. Shuffling marker positions further worsened the accuracy, and almost all missing data point assigned to the majority class.\\
FILLIN performances on rice follow the same pattern and worsen substantially when markers are shuffled. This result confirmed that imputation methods specifically designed for ordered markers are not an option for species lacking a reference genome.
